---
title: "Plots 1/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(kdensity)
library(car)
library(tidyr)
load("data/asv112_longer.RData") #dat_longer
load("data/asv112_summaries.RData") #dat_summaries
dat_summaries_eps <- dat_summaries
dat_summaries_eps$cumRelAbund[which(dat_summaries_eps$cumRelAbund == 0)] <- .Machine$double.eps
```
## Density estimates

### Cumulative relative abundance

```{r densities, echo=F}
dat_summaries %>% filter(cumRelAbund > 0) %>%
  ggplot(aes(x = log10(cumRelAbund), fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black')
  geom_density(alpha = 0.7) +
  geom_vline(aes(xintercept = log10(1e-2)), 
             color = 'black', 
             linetype = 'dashed') + 
  labs(x = 'log10 Cumulative Relative Abundance (zeros removed)')
dat_summaries_eps %>% ggplot(aes(x = log10(cumRelAbund), fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black')
  geom_density(alpha = 0.7) +
  geom_vline(aes(xintercept = log10(1e-2)), 
             color = 'black', 
             linetype = 'dashed') + 
  labs(x = 'log10 Cumulative Relative Abundance (zeros replaced by machine epsilon)')
dat_summaries %>% filter(cumRelAbund > 0) %>%
  ggplot(aes(x = logit(cumRelAbund), fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black')
  geom_density(alpha = 0.7) +
  geom_vline(aes(xintercept = logit(1e-2)), 
           color = 'black', 
           linetype = 'dashed') + 
  labs(x = 'logit Cumulative Relative Abundance (zeros removed)')
dat_summaries %>%
  ggplot(aes(x = cumRelAbund, fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black') +
  geom_density(alpha = 0.7) +
  labs(x = 'Cumulative Relative Abundance') +
  xlim(0, 0.02) + 
  geom_vline(aes(xintercept = 1e-2), 
             color = 'black', 
             linetype = 'dashed')
dat_summaries %>% filter(cumRelAbund > 0) %>%
  ggplot(aes(x = cumRelAbund, fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black')
  geom_density(alpha = 0.7) +
  geom_vline(aes(xintercept = 1e-2), 
           color = 'black', 
           linetype = 'dashed') + 
  labs(x = 'Cumulative Relative Abundance (zeros removed)') +
  xlim(0.0, 0.02)
```

### Total count 
```{r totalcount, echo=FALSE}
dat_summaries %>% filter(totalCount > 0) %>%
  ggplot(aes(x = log10(totalCount), fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black')
  geom_density(alpha = 0.7) +
  geom_vline(aes(xintercept = log10(100)), 
             color = 'black', 
             linetype = 'dashed') + 
  labs(x = 'log10 Total Count (zeros removed)')
dat_summaries %>%
  ggplot(aes(x = totalCount, fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black')
  geom_density(alpha = 0.7) +
  labs(x = 'Total Count') +
  geom_vline(aes(xintercept = 100), 
             color = 'black', 
             linetype = 'dashed') + 
  xlim(0, 1000)
dat_summaries %>% filter(totalCount > 0) %>%
  ggplot(aes(x = totalCount, fill = geno)) +
  # geom_histogram(alpha = 0.7,
  #                position = 'identity',
  #                aes(y = ..density..),
  #                color = 'black')
  geom_density(alpha = 0.7) +
  geom_vline(aes(xintercept = 100), 
           color = 'black', 
           linetype = 'dashed') + 
  labs(x = 'Total Count (zeros removed)') + 
  xlim(0,1000)
```

## Bar plots of total presence 
```{r bar plots, echo=F}
bar <- ggplot(dat_summaries, aes(x = genoPresence)) +
  geom_bar() +
  facet_wrap(.~geno) +
  labs(x = 'number of samples in which ASV is present', y = 'number of ASVs')
bar
```

## Notes
The histogram and KDE are probably boundary effected near zero for the untransformed CRA and total count, i.e. underestimate density at zero and very near zero. I know in theory how to mitigate this effect, but I'm not sure it's worth it for current purposes (geom_density calls density(), which doesn't have the fixes built-in). 

```{r boundarybias}

#large bandwidth
dat_summaries %>% filter(geno == 'Grassl') %>% 
  ggplot(aes(x = cumRelAbund, fill = geno)) +
  geom_histogram(alpha = 0.9,
                 position = 'identity',
                 aes(y = ..density..),
                 color = 'black',
                 binwidth = 0.0001) +
  geom_density(alpha = 0.5, adjust = 30) + 
  labs(x = 'CRA near 0, large bw') +
  xlim(0, 0.002) 

#small bandwidth
dat_summaries %>% filter(geno == 'Grassl') %>% 
  ggplot(aes(x = cumRelAbund, fill = geno)) +
  geom_histogram(alpha = 0.9,
                 position = 'identity',
                 aes(y = ..density..),
                 color = 'black',
                 binwidth = 0.0001) +
  geom_density(alpha = 0.5, adjust = 1/2) + 
  labs(x = 'CRA near 0, with histogram') +
  xlim(0, 0.002) 

grassl <- dat_summaries %>% filter(geno == 'Grassl')
nrow(grassl %>% filter(cumRelAbund == 0)) / nrow(grassl)
nrow(grassl %>% filter(between(cumRelAbund, .Machine$double.eps, 0.0001)))/nrow(grassl)
nrow(grassl %>% filter(between(cumRelAbund, 0.0001, 0.0002)))/nrow(grassl)
nrow(grassl %>% filter(between(cumRelAbund, 0.0002, 0.0003)))/nrow(grassl)
nrow(grassl %>% filter(between(cumRelAbund, 0.0003, 0.0004)))/nrow(grassl)
nrow(grassl %>% filter(between(cumRelAbund, 0.0004, 0.0005)))/nrow(grassl)
```

